Supermartingales and Termination Analysis of Statistical PCF

### 2021-09-19 LICS21/intro.tex
Removed \onecolumn and \twocolumn - seems to compile now

- Use {\tt citep} to render ``(\emph{author}, \emph{year})'' as opposed to ``\emph{author} (\emph{year})''}

@@

Basic (non-random) lambda calculus has the very useful property of confluence, which implies (among other things) that even if execution of a program starts in a different order, it will still reach the same normal form eventually (assuming it does reach a normal form). 
Probabilistic lambda calculus does not have this property, because random choices may be duplicated, and evaluating the same subterm multiple times can yield different results. 
However, with a restricted set of reduction strategies, confluence may be regained. 
We introduce a novel addressing scheme for the possible random choices in a program's execution, which ensures that the same random choices are taken at corresponding positions in alternative reduction sequences, so that the same eventual result can be reached. 
This is then used to prove yet another extension to the ranking function theorem, that (\cref{thm:confluent ranking}) ranking functions may be defined with respect to alternative reduction strategies (which in some cases may lead to a considerably simpler execution and ranking function), and (\cref{cor:Reduction strategy independence}) rankability in this sense still imples almost sure termination. 
The confluent trace semantics is of wider interest, and has other possible applications as well, for example in Bayesian inference algorithms.

>>>>

- Use {\tt citep} to render ``(\emph{author}, \emph{year})''; default {\tt cite} is ``\emph{author} (\emph{year})''}

@@

Basic (non-random) lambda calculus has the very useful property of confluence, which implies (among other things) that even if execution of a program starts in a different order, it will still reach the same normal form eventually (assuming it does reach a normal form). 
Probabilistic lambda calculus does not have this property, because random choices may be duplicated, and evaluating the same subterm multiple times can yield different results. 
However, with a restricted set of reduction strategies, confluence may be regained. 
We introduce a novel addressing scheme for the possible random choices in a program's execution, which ensures that the same random choices are taken at corresponding positions in alternative reduction sequences, so that the same eventual result can be reached. 
This is then used to prove yet another extension to the ranking function theorem, that (\cref{thm:confluent ranking}) \lo{Preceding reference generated is wrong; it should be a theorem} ranking functions may be defined with respect to alternative reduction strategies (which in some cases may lead to a considerably simpler execution and ranking function), and (\cref{cor:Reduction strategy independence}) rankability in this sense still imples almost sure termination. 
The confluent trace semantics is of wider interest, and has other possible applications as well, for example in Bayesian inference algorithms.
